---
output:
  html_document:
    df_print: paged
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, warning = FALSE, message = FALSE)
options(knitr.kable.NA = '')
```

<hr>

```{r load results, message=FALSE, results='hide'}
source("/Users/cjcampbell/Pd_DetectionMigratoryBats/R/00_Setup.R")
source("/Users/cjcampbell/Pd_DetectionMigratoryBats/R/01_load_qPCR_results.R")
```

##### Data from Table 1: Individuals of each species sampled by state and range of years and month of year sampled.

```{r table1, warning=FALSE}
sum_tab <- mdf %>% 
  dplyr::filter(!is.na(Well), !is.na(Spp)) %>% 
  dplyr::group_by(Spp, Source_St) %>% 
  dplyr::mutate(
    Year = as.numeric(Year),
    Month = as.numeric(Month),
    ) %>% 
  dplyr::summarise(
    n_indiv = length(unique(AL.CODE)) ,
    min_yr = abs( min( Year,  na.rm = TRUE) ),
    max_yr = abs( max( Year,  na.rm = TRUE) ),
    min_mo = abs( min( Month, na.rm = TRUE) ),
    max_mo = abs( max( Month, na.rm = TRUE) )
  ) %>% 
  dplyr::mutate(
    range_year  = if_else(min_yr == max_yr, paste(min_yr), paste(min_yr, " - ", max_yr) ),
    range_month = if_else(min_mo == max_mo, paste(min_mo), paste(min_mo, " - ", max_mo) )
    ) %>% 
  ungroup() %>% 
  dplyr::select(Spp, Source_St, range_year, range_month, n_indiv) %>% 
  arrange(Spp, Source_St, desc(n_indiv))

sum_tab %>% kable()
```
<hr>

##### Data included in Table 2: Number of individuals sampled by each sampling method. Note that individuals sampled by multiple methods (e.g., “tissue and skin swab”) correspond with multiple samples.

```{r table2}

mdf %>% 
  dplyr::filter(!is.na(Well), !is.na(Spp)) %>% 
  dplyr::group_by(AL.CODE, Spp, sampleType, `Sampling Method`) %>% 
  dplyr::summarise(n = n()) %>% 
  ungroup() %>% 
  dplyr::select(-n) %>% 
  dplyr::mutate(
    samplingMethod = case_when(
      grepl("Turbine", `Sampling Method`, ignore.case = T) ~ "carcass",
      grepl("live", `Sampling Method`, ignore.case = T)    ~ "live",
      grepl("netting", `Sampling Method`, ignore.case = T)    ~ "live",
      TRUE ~ "Unknown"
    )
  ) %>% 
  dplyr::group_by(AL.CODE, Spp, sampleType, samplingMethod) %>% 
  dplyr::summarise(n = n()) %>% 
  pivot_wider(names_from = sampleType, values_from = n) %>% 
  dplyr::mutate(
    sampled = case_when(
      extractedDNA==1&swab_skin==1&swab_fur==1 ~ "All",
      extractedDNA==1&is.na(swab_skin)&swab_fur==1 ~ "Tissue and fur swab",
      extractedDNA==1&swab_skin==1&is.na(swab_fur) ~ "Tissue and skin swab",
      extractedDNA==1&is.na(swab_skin)&is.na(swab_fur) ~ "Tissue only",
      is.na(extractedDNA)&swab_skin==1&swab_fur==1 ~ "Skin and fur swabs",
      is.na(extractedDNA)&swab_skin==1&is.na(swab_fur) ~ "Skin swab only",
      is.na(extractedDNA)&is.na(swab_skin)&swab_fur==1 ~ "Fur swab only",
  )) %>% 
  dplyr::group_by(Spp, samplingMethod, sampled) %>% 
  dplyr::summarise(n = n()) %>% 
  ungroup() %>% 
  pivot_wider(values_from = n, names_from = sampled) %>% 
  kable()
```

<hr>

##### Data from table 3: Table 3: Sampling details for Pd-positive individuals. One individual represented here was discarded due to irregular curve results (i.e., Cт very low).

```{r table3 }

has_pos_ID <- mdf %>% 
  filter(Cт != "Undetermined") %>% 
  dplyr::select(AL.CODE) %>% 
  na.omit() %>% 
  distinct %>% 
  unlist()

pos_results_table <- mdf %>% 
  filter(AL.CODE %in% has_pos_ID) %>% 
  dplyr::mutate(Cт = round(as.numeric(Cт), digits = 2)) %>% 
  arrange(Spp, AL.CODE, desc(Cт), sampleType) %>% 
  group_by(Spp, Source_St, Year, Month, Day, AL.CODE, sampleType) %>% 
  summarise(Ct_vector=paste(Cт, collapse=", ")) %>% 
  dplyr::select(Spp, Source_St, Year, Month, Day, sampleType, Ct_vector)

pos_results_table %>% 
  as.data.frame %>% 
  dplyr::select(-AL.CODE) %>% 
  dplyr::rename(
    "Species" = Spp, 
    "State" = Source_St, 
    "Sample type" = sampleType,
    "Cт results" = Ct_vector
    ) %>% 
  dplyr::mutate(
    Species = recode(
      Species, 
      "LACI" = "Lasiurus cinereus", 
      "LABO" = "Lasiurus borealis", 
      "LANO" = "Lasionycteris noctivagans"
      ),
    `Sample type` = recode(
      `Sample type`,
      extractedDNA = "Extracted DNA",
      swab_fur = "Fur swab",
      swab_skin = "Skin swab"
      )
    ) %>% 
  knitr::kable(align = c(rep("l", 2), rep("r", 3), "l", "r"), padding = 2) %>% 
    column_spec(1, width = "2in") %>% 
    column_spec(2, width = "0.4in") %>% 
    column_spec(3, width = "0.4in") %>% 
    column_spec(4, width = "0.4in") %>% 
    column_spec(5, width = "0.75in") %>% 
    column_spec(6, width = "1.25in") %>% 
    column_spec(7, width = "2in")
```

<hr>

```{r SI_fig_qPCR results}
if(!exists("results_files")) stop("object results_files is missing! Make sure you load the data.")
suppressWarnings({ suppressMessages({
  qPCR_out <- lapply(results_files, function(file){
    
    data_main <- readxl::read_excel(file, range = "A8:AB104", sheet = 2)
    
    data_header <- readxl::read_excel(file, range = "A1:B6", col_names = FALSE, sheet = 2) %>% 
      t() %>% as.data.frame(stringsAsFactors = FALSE)
    colnames(data_header) <- data_header[1,]
    data_header <- data_header[-1,]
    
    data_footer <- readxl::read_excel(file, range = "A106:B109", col_names = FALSE, sheet = 2) %>% 
      t() %>% as.data.frame(stringsAsFactors = FALSE)
    colnames(data_footer) <- data_footer[1,]
    data_footer <- data_footer[-1,]
    
    maindat <- purrr::reduce(list(data_main, data_header, data_footer), cbind)
   
    raw_data_main <- readxl::read_excel(file, range = "A8:E3848", sheet = 5)
    
    raw_data_header <- readxl::read_excel(file, range = "A1:B6", col_names = FALSE, sheet = 5) %>% 
      t() %>% as.data.frame(stringsAsFactors = FALSE)
    colnames(raw_data_header) <- raw_data_header[1,]
    raw_data_header <- raw_data_header[-1,]
  
    rawdat <- purrr::reduce(list(raw_data_main, raw_data_header), cbind)
  
    return( left_join(rawdat, maindat) )
  }) %>% 
    plyr::ldply() %>% 
    mutate(
      sample_type = if_else( grepl( "dilution", `Sample Name`), "dilution",
                             if_else( grepl("Pd_PC",  `Sample Name`), "control_pos",
                                      if_else( grepl("NC",  `Sample Name`), "control_neg",
                                               "sample"
                                      )))
      
    ) %>% 
    dplyr::mutate( plateName = 
                     tools::file_path_sans_ext( 
                       gsub( "D:Applied BiosystemsStepOne Software v2.2.2experimentsPdCaitlin_Pd_", "", 
                             gsub("\\\\", "", `Experiment File Name`) ) ) )
})})

p <- qPCR_out %>% 
  filter(sample_type != "dilution") %>% 
  ggplot() +
  geom_line(
    aes(
      x = Cycle, y = ΔRn, 
      group = interaction(Well, `Sample Name`, plateName),
      color = sample_type 
    ),
    size = 0.4
  ) +
  geom_hline(
    aes(yintercept = `Ct Threshold`, group = interaction(Well, `Sample Name`, plateName)),
    linetype = 2
    ) +
  facet_wrap(~plateName, scales= "free", ncol = 3) +
  scale_y_log10(limits = c(1e-4, 1e1)) +
  ggplot2::scale_colour_viridis_d(
    direction = -1 ,
    name = "Sample type",
    labels = c("Neg. control", "Pos. control", "Samples")
    ) +
  theme_bw() +
  theme(
    axis.line = element_line(colour = "black"),
    panel.grid.major = element_blank(),
    panel.grid.minor = element_blank(),
    panel.border = element_blank(),
    panel.background = element_blank(),
    strip.background = element_blank(),
    legend.position = c(0.85, 0.05)
  )
p
```


##### SI Figure 1: Response curves from qPCR analyses. Each panel represents a distinct plate run, with the date and sequence of each run identified in the panel title. Dotted lines indicate run-specific thresholds. Control positives (teal) and negatives (yellow) all performed as expected.
